MT @HanBaldwin: Fightin OOVs in German \#twitter

### 2013-07-31 darmstadt_gscl_2013/article_twitter_normalization.tex
statistics updated

  This article gives an overview of existing approaches to the problem
  of out-of-vocabulary (OOV) tokens and noisiness phenomena in natural
  language texts. These approaches are classified with regard to the
  size of text spans and knowledge inference mechanisms which they
  rely on in their work. Additionally to that, we conduct quantitative
  and qualitative analyses of unknown words in German Twitter
  messages, in order to see how relevant the OOV and text
  normalization problems are for this particular kind of micro-texts
  and what the characteristics of these OOVs are. In a concluding
  step, we present a set of ad-hoc techniques which are supposed to
  tackle some of the most prominent disturbing effects found during
  these analyses and show how this set of techniques helps us lower
  the average rate of out-of-vocabulary tokens in Twitter messages and
  how this lower OOV-rate helps improve the quality of automatic
  Part-of-speech tagging.

@@

So, Kukich (\cite{kukich}), for example, divided NTN techniques into
six classes:
\begin{enumerate}
\item minimum edit distance techniques;
\item similarity key techniques;
\item rule-based techniques;
\item \textit{n}-gram based techniques;
\item probabilistic techniques;
\item neural nets.
\end{enumerate}
Kobus (\cite{kobus}), on the contrary, referred to NTN methods as
\textit{metaphors} and split them into the following three groups:
\begin{enumerate}
\item ``spell checking'' metaphor;
\item ``translation'' metaphor;
\item ``speech recognition'' metaphor.
\end{enumerate}

@@

The reason for this confusion is the fact that the above
classifications both rely on several independent criteria at the same
time though each of these criteria characterizes an NTN system from a
different point of view. As a consequence, unambiguous assignment of
an NTN approach to one particular class often becomes impossible. In
order to avoid this, we instead suggest using separate classifications
for each criterion without mixing them.

One of such criteria which in our opinion would be worth a separate
classification is \emph{segmentation level} which is used
\begin{inparaenum}[\itshape a\upshape)]
\item to infer in-vocabulary (IV) equivalents for OOV
  tokens\label{candidate:inferring} and
\item to choose the most probable variant\label{candidate:selection}
  among multiple possible ones\footnote{For the cases, when segments
    of different lengths are used for tasks \ref{candidate:inferring}
    and \ref{candidate:selection}, we note it explicitly in our
    classification on which segmentation length each of these subtasks
    relies.}.
\end{inparaenum} For this criterion, we propose division into the following
classes:
\begin{enumerate}
  \item graphematic\footnote{Depending on whether phonetical
    information is involved or not at this level, this class could be
    further divided into a phonographematic and purely graphematic
    subclasses.};
  \item lexical;
  \item phrasal.
\end{enumerate}
Each broader level of this hierarchy is also supposed to either
incorporate or ignore information provided by its narrower
subsegments. In this way, we only need to mention one (the broadest)
hierarchical class for the cases when multiple segmentation levels are
involved by some techniques.

@@

The earlier works on NTN commonly used either purely graphematic or
phonographematic levels of segmentation for deriving normalization
variants of incorrectly spelled words. To purely graphematic systems
belong methods suggested Brill and Moore (\cite{brill}), Sproat et
al. (\cite{sproat}), and Clark (\cite{alexander-clark}). As
phonographematic approaches one could regard works done by Toutanova
and Moore (\cite{toutanova}), Choudhury et al. (\cite{choudhury}),
Cook and Stevenson (\cite{cook}) etc. With regard to the type of
information inference, practically all of these methods were
supervised with the exception of Cook and Stevenson (\cite{cook}) who
claimed to uns an unsupervised technique.

Starting from the second half of the 2000s, the raising influence and
improved quality of machine translation tools lead to the development
of NTN technologies which used broader levels of segmentation. In
\cite{aw}, Aw et al. suggested a supervised statistical system for
normalization of SMS-messages which operated on automatically aligned
phrases. A few years later, Clark and Araki (\cite{clark-araki})
described a purely rule-based method which used mappings of
non-standard words and phrases to their corresponding standard
language forms.

As noted by Kobus et al. (\cite{kobus}), NTN methods relying on either
graphematic or phrasal segments usually revealed complementary
strengths and weaknesses. This notion led NLP scientists to the idea
that by incorporating multiple levels of the language into one NTN
system the total performace of the system would improve as different
sources of information would benefit from each other. As a consequence
of this, a wealth of combined techniques emerged in the past few
years. Among these we should especially mention the works by Kobus et
al. (\cite{kobus}), Kaufmann (\cite{kaufmann}), Han and Baldwin
(\cite{han}), and Oliva et al. (\cite{oliva}). These systems used the
whole range of segmentation levels from phonographematatic to the
phrasal one, and also different knowledge inference techniques were
applied for different levels.

@@

It should however be noted that almost all of the above methods mainly
concentrated on only English text data. A few exceptions from that are
works done by Beaufort et al. (\cite{beaufort}) for French and Oliva
et al. (\cite{oliva}) for Spanish. To find out which peculiarities of
ill-formed words are characteristic for German, we will perform a
quantitive and qualitative analysis of unknown words in German Twitter
in the next Section to see what kind of NTN techniques would be more
suitable for handling ill-formed words there.

@@

In order to estimate the percentage of unknown words in Twitter
messages, we randomly selected 10,000 tweets from a previously
collected corpus, split them into sentences and tokenized using social
media aware tokenizer by Christopher Potts
\footnote{\url{http://sentiment.christopherpotts.net/code-data/happyfuntokenizing.py}}.
After skipping all words which did not contain any alphabetic
character or consisted only of a single letter, we
obtained a list of 129,146 tokens. As reference systems for dictionary
lookup we used open source spell checking program
\texttt{hunspell}\footnote{Ispell Version 3.2.06 (Hunspell Version
  1.3.2); dictionary de\_DE.} and publicly available part-of-speech
tagger \texttt{TreeTagger}\footnote{Version 3.2 with German parameter
  file UTF-8.} (Schmid, \cite{schmid}).

@@

Here once again the question of classification arose but this time
with regard to the reasons, why different tokens could have been
omitted from corresponding applications' dictionaries. In this
respect, division into following groups seemed to be appropriate for
us:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionary
    (MRD)}. To this group we counted words of basic vocabulary which
    did not get into applications' MRD either because they supposedly
    were rare or because they did not exist at the time when
    dictionaries were created. Another reason for inclusion in this
    type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are usually deliberately not included in MRDs
    due to impossibility to fully cover these classes;\label{dict}
  \item \textbf{Unintended sloppiness of users' input}. In the scope
    of this group we considered unintended typos including unexpected
    truncation of words at the end of Twitter messages\footnote{As is
      generally known, Twitter imposes a strong restriction on the
      length of posted messages which can be no longer than 140
      characters. Upon exceeding this length, tweets get automatically
      truncated to maximal allowed length.};\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language but were perfectly valid terms
    in the domain of web discourse and particularly Twitter
    communication.\label{style}
\end{enumerate}

In order to see how out-of-vocabulary words were distributed among and
within these 3 major groups, we manually analyzed all OOV tokens which
appeared in text more than once and also looked at 1,000 randomly
selected hapax legomena. The results of these estimations are shown
and explained below.

We subdivided group \ref{dict} into the following subgroups:
\begin{enumerate}
\item regular German words, e.g. \textit{aufm},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{geskypt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexicons and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrarily to a certain degree and
involves different linguistic criteria simultaneously, the underlying
notion for it was simple. Valid words could have been omitted from an
MRD either due to the limitation of developers' capacities (group
\ref{regular}), active word formation or lexical productivity of the
language (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

In Table \ref{table:mrd} percentage figures for each of the above
subgroups are shown. We have considered OOV-distribution for both
\texttt{hunspell} and \texttt{TreeTagger}. For both of them we
estimated the percentage of particular subclass with regard to the
total number of occurrences of OOV-tokens (column ``\totaloov{}'') as
well as with regard to their percentage in the list of unique OOVs
(column ``\uniqoov{}'') without taking into account their frequencies.

@@

Similarly to class \ref{dict} we subdivided the group of misspellings
into the following subgroups:
\begin{enumerate}
  \item insertion;
  \item deletion;
  \item substitution;
\end{enumerate}
according to the kind of operation which led to a particular spelling
mistake. The statistics on distribution of these subgroups is shown in
Table \ref{table:spell}.
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``unintended sloppiness of users' input''\label{table:spell}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    insertion &  &  &  & \\
    deletion &  &  &  & \\
    substitution &  &  &  & \\\hline
    {\bfseries total} &  &  &  & \\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

>>>>

  This article gives an overview of existing approaches to the problem
  of out-of-vocabulary (OOV) tokens and noisiness phenomena in natural
  language texts. These approaches are classified with regard to the
  size of text spans and knowledge inference mechanisms which they
  rely on in their work. Additionally to that, we conduct quantitative
  and qualitative analyses of unknown words in German Twitter
  messages, in order to see how relevant the OOV and text
  normalization problems are for this particular kind of micro-texts
  and what the characteristics of these OOVs are. In a concluding
  step, we present a set of ad-hoc techniques which are supposed to
  tackle some of the most prominent disturbing effects found during
  these analyses and show how this set of techniques helps us lower
  the average rate of out-of-vocabulary tokens in Twitter messages and
  how this lower OOV-rate helps improve the quality of automatic
  part-of-speech tagging.

@@

Kukich (\cite{kukich}), for example, divided NTN techniques into six
classes:
\begin{enumerate}
\item minimum edit distance techniques;
\item similarity key techniques;
\item rule-based techniques;
\item \textit{n}-gram based techniques;
\item probabilistic techniques;
\item neural nets.
\end{enumerate}
Kobus (\cite{kobus}), on the contrary, referred to NTN methods as
\textit{metaphors} and split them into the following three groups:
\begin{enumerate}
\item ``spell checking'' metaphor;
\item ``translation'' metaphor;
\item ``speech recognition'' metaphor.
\end{enumerate}

@@

The reason for this confusion is the fact that the above
classifications both rely on several independent criteria at the same
time though each of these criteria characterizes an NTN system from a
different point of view. As a consequence, unambiguous assignment of
an NTN approach to one particular class often becomes impossible. In
order to avoid this, we instead suggest using separate classifications
for each criterion which might characterize an NTN technique.

One of such criteria which in our opinion would be worth a separate
classification is \emph{segmentation level} which is used
\begin{inparaenum}[\itshape a\upshape)]
\item to infer in-vocabulary (IV) equivalents for OOV
  tokens\label{candidate:inferring} and
\item to choose the most probable variant\label{candidate:selection}
  among multiple possible suggestions\footnote{For the cases, when
    segments of different lengths are used for tasks
    \ref{candidate:inferring} and \ref{candidate:selection}, we note
    it explicitly in our classification on which segmentation length
    each of these subtasks relies.}.
\end{inparaenum} For this criterion, we propose division into the following
classes:
\begin{enumerate}
  \item graphematic\footnote{Depending on whether phonetical
    information is involved or not at this level, this class could be
    further divided into a phonographematic and purely graphematic
    subclasses.};
  \item lexical;
  \item phrasal.
\end{enumerate}
Each broader level of this hierarchy is supposed to either incorporate
or ignore information provided by its narrower subsegments. In this
way, we only need to mention one (the broadest) hierarchical class for
the cases when multiple segmentation levels are involved by some
techniques.

@@

The earlier works on NTN commonly relied on either purely graphematic
or phonographematic levels of segmentation for deriving normalization
variants of incorrectly spelled words. To purely graphematic systems
belong methods suggested Brill and Moore (\cite{brill}), Sproat et
al. (\cite{sproat}), and Clark (\cite{alexander-clark}). As
phonographematic approaches one could regard works done by Toutanova
and Moore (\cite{toutanova}), Choudhury et al. (\cite{choudhury}),
Cook and Stevenson (\cite{cook}) etc. With regard to the type of
information inference, practically all of these methods were
supervised with the exception of Cook and Stevenson (\cite{cook}) who
claimed to uns an unsupervised technique.

Starting from the second half of the 2000s, the raising influence and
improved quality of machine translation tools lead to the development
of NTN technologies which used broader levels of segmentation. In
\cite{aw}, Aw et al. suggested a supervised statistical system for
normalization of SMS-messages which operated on automatically aligned
phrases A few years later, Clark and Araki (\cite{clark-araki})
described a purely rule-based method which used mappings of
non-standard words and phrases to their corresponding standard
language forms.

As noted by Kobus et al. (\cite{kobus}), NTN methods relying on either
graphematic or phrasal segments usually revealed complementary
strengths and weaknesses. This notion led NLP scientists to the idea
that by incorporating multiple levels of the language into one NTN
system the total performace of the whole system would improve as
different sources of information would benefit from each other. As a
consequence of this, a wealth of combined techniques emerged in the
past few years. Among these we should especially mention works by
Kobus et al. (\cite{kobus}), Kaufmann (\cite{kaufmann}), Han and
Baldwin (\cite{han}), and Oliva et al. (\cite{oliva}). The majority of
these systems used the whole range of segmentation levels from
phonographematatic to phrasal one, and in many cases they also applied
different knowledge inference mechanisms to different levels of the
language.

@@

It should however be noted that almost all of the above methods mainly
concentrated on only English data. A few exceptions from that are
approaches suggested by Beaufort et al. (\cite{beaufort}) for French
and Oliva et al. (\cite{oliva}) for Spanish. To find out which
peculiarities of ill-formed words are characteristic for German, we
will perform a quantitive and qualitative analysis of unknown words in
German Twitter in the next Section in order to see what kind of NTN
techniques would be more suitable for handling ill-formed words there.

@@

In order to estimate the percentage of unknown words in Twitter
messages, we randomly selected 10,000 tweets from a previously
collected corpus, split them into sentences and tokenized using social
media-aware tokenizer by Christopher Potts
\footnote{\url{http://sentiment.christopherpotts.net/code-data/happyfuntokenizing.py}}.
After skipping all words which did not contain any alphabetic
character in them or consisted only of a single letter, we obtained a
list of 129,146 tokens. As reference systems for dictionary lookup we
used open source spell checking program
\texttt{hunspell}\footnote{Ispell Version 3.2.06 (Hunspell Version
  1.3.2); dictionary de\_DE.} and publicly available part-of-speech
tagger \texttt{TreeTagger}\footnote{Version 3.2 with German parameter
  file UTF-8.} (Schmid, \cite{schmid}).

@@

Here once again the question of classification arose. This time with
regard to the reasons, why different tokens could have been omitted
from corresponding applications' dictionaries. In this respect,
division into following groups seemed to be appropriate for us:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionary
    (MRD)}. To this group we counted words of basic vocabulary which
    did not get into applications' MRD either because they supposedly
    were rare or because they did not exist at the time when
    dictionaries were created. Another reason for the inclusion in
    this type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are often omitted from MRDs due to
    impossibility to fully cover these classes;\label{dict}
  \item \textbf{Unintended sloppiness of users' input}. In the scope
    of this second group, we considered unintended typos including
    unexpected truncation of words at the end of Twitter
    messages\footnote{As is generally known, Twitter imposes a strong
      restriction on the length of posted messages which can be no
      longer than 140 characters. Upon exceeding this length, tweets
      get automatically truncated to maximal allowed
      length.};\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language texts but were perfectly valid
    terms in the domain of web discourse and Twitter communication in
    particular.\label{style}
\end{enumerate}

In order to see how detected out-of-vocabulary words were distributed
among and within these 3 major groups, we manually analyzed all OOV
tokens which appeared in text more than once and also looked at 1,000
randomly selected hapax legomena. The results of these estimations are
shown and explained below.

We subdivided group \ref{dict} into the following subgroups:
\begin{enumerate}
\item regular German words, e.g. \textit{aufm},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{gegoogelt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexicons and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrarily to a certain degree and
involves different linguistic criteria simultaneously, the underlying
notion for it is simple. Valid words could have been omitted from an
MRD either due to the limitation of developers' capacities (group
\ref{regular}), active word formation or lexical productivity of the
language (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

In Table \ref{table:mrd}, percentage figures for each of the above
subgroups are shown. We have considered OOV-distribution for both
\texttt{hunspell} and \texttt{TreeTagger}. For each of them, we
estimated the percentage of a particular subclass with regard to the
total number of occurrences of OOV-tokens (column ``\totaloov{}'') as
well as with regard to their percentage rate in the list of only
unique OOVs (column ``\uniqoov{}''), i.e. disregarding their
frequencies.

@@

Similarly to class \ref{dict}, we subdivided the group of misspellings
into the following subgroups:
\begin{enumerate}
  \item insertion;
  \item deletion;
  \item substitution;
\end{enumerate}
according to the kind of operation which led to a particular spelling
mistake. The statistics on distribution of these subgroups is shown in
Table \ref{table:spell}.
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``unintended sloppiness of users' input''\label{table:spell}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    insertion &  &  &  & \\
    deletion &  &  &  & \\
    substitution &  &  &  & \\\hline
    {\bfseries total} &  &  &  & \\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}
